|

### 2020-05-13 paper/Manuscript.Rmd
Updat Manuscript
Added some info on data sharing and personal/pseudononymized data

One important impediment to comprehensive citation is the fact that print journals operate with space constraints. Print journals often discourage comprehensive citation, either actively, or passively by including the reference list in the manuscript word count. Researchers can overcome this impediment by preparing two versions of the manuscript: One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations. The print version should reference the online version, so interested readers can find the comprehensive reference list. Step 20 of the WORCS procedure suggests uploading the online version to a preprint server. This is important because most preprint servers - [at least the ones hosted by OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar. This means that authors will receive credit for cited work; even if they are cited only in the online version. Moreover, preprint servers ensure that the online version will have a persistent DOI, and will remain reliably accessible, just like the print version.

@@

Data sharing is important for computational reproducibility and secondary analysis. Computational reproducibility means that a third party can exactly recreate the results from the raw data, using the published analysis code. Secondary analysis means that a third party can conduct sensitivity analyses, explore alternative explanations, or even use existing data to answer a different research question. From an open science perspective, data sharing is always desirable. From a practical point of view, it is not always possible. Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns. For example, the GDPR prohibits storing 'personal data' (information which can identify a natural person whether directly or indirectly) on a server outside the EU. Different rules may apply to anonymized or pseudonimized data. Before sharing any data, it is recommended to obtain approval from an internal (ethical) review board, guidance from Research Data Management Support, and informed consent from participants.

<!-- LB: As I reader, I would prefer to hear about data sharing options before the synthetic data part (which is only relevant to those who cannot share their data). E.g. can I indicate a point in my workflow (e.g. after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it. Any data processing steps should be documented in the code, instead of applied to the raw data. This way, the code will be a pipeline from the raw data to the published results.

>>>>

One important impediment to comprehensive citation is the fact that print journals operate with space constraints. Print journals often discourage comprehensive citation, either actively, or passively by including the reference list in the manuscript word count. Researchers can overcome this impediment by preparing two versions of the manuscript: One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations. The print version should reference the online version, so interested readers can find the comprehensive reference list. Step 20 of the WORCS procedure suggests uploading the online version to a preprint server. This is important because most preprint servers - [at least the ones hosted by the Open Science Framework OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar. This means that authors will receive credit for cited work; even if they are cited only in the online version. Moreover, preprint servers ensure that the online version will have a persistent DOI, and will remain reliably accessible, just like the print version.

@@

Data sharing is important for computational reproducibility and secondary analysis. Computational reproducibility means that a third party can exactly recreate the results from the raw data, using the published analysis code. Secondary analysis means that a third party can conduct sensitivity analyses, explore alternative explanations, or even use existing data to answer a different research question. From an open science perspective, data sharing is always desirable. From a practical point of view, it is not always possible. Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns. When working with data from human participants it is important to delete any sensitive personal data and contact information as soon as data collection has finished. The resulting pseudononimized data can often be shared if informed consent was given. Nevertheles, it is important to note that the GDPR in europe prohibits storing 'personal data' (information which can identify a natural person whether directly or indirectly) on a server outside the EU, unless they offer an "adequate level of protection". Although different rules may apply to pseudonimized data, there are repositories with EU servers such as the Open Science Framework and universities might have their own repositories that are complient with local legistation. Before sharing any human participant data, it is recommended to obtain approval from an internal (ethical) review board, guidance from Research Data Management Support, and informed consent from participants.

<!-- LB: As I reader, I would prefer to hear about data sharing options before the synthetic data part (which is only relevant to those who cannot share their data). E.g. can I indicate a point in my workflow (e.g. after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->When data can be shared, it is important to version control the pseudonymized <!-- MS: it is almost impossible to anonymize data, so it's preferred to use pseudononymized ---> raw data upon collection, and avoid changing it. Any data processing steps should be documented in the code, instead of applied to the raw data. This way, the code will be a pipeline from the raw data to the published results.
